---
title: "Perceptron Example"
author: "Jeffrey Norton"
date: "January 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# McCullogh-Pitts Neurons

There are several references which are helpful in understanding McCullogh-Pitts neurons,
but [this one](https://www.lri.fr/~marc/EEAAX/Neurones/tutorial/mcpits/html/index.html)
was helpful.  The key formula is given by Hinton in his PDF notes.

1) Compute a weighted sum plus a bias.
$z = b + \sum_i x_i w_i$.
2) Output 1 if $z > 0$.  $y = 1 $ if $y \gt 0$, $0$ otherwise

## Example
The inputs $x$ consist of excitatory inputs with values ${0, 1}$,
```{r, echo=TRUE, eval=TRUE, cache=TRUE}
n=10
(x_excite=sample(c(0,1), n, replace=TRUE))
(w_excite=runif(n, min=0, max=1))
```
inhibitory inputs with values ${0, 1}$,
```{r, echo=TRUE, eval=TRUE, cache=TRUE}
m=8
(x_inhibit=sample(c(-1,0), m, replace=TRUE))
(w_inhibit=runif(m, min=0, max=1))
```
and the threshold of $0$ as described above
```{r, echo=TRUE, eval=TRUE, cache=TRUE}
(u = 0)
```
Calculate the weighted sum as shown above with bias of $0.5$.
```{r, echo=TRUE, eval=TRUE, cache=TRUE}
b = 0.5
(z = b + sum(x_excite, w_excite) + sum(x_inhibit, w_inhibit))
```
Calculate the neuron output of $y$ using the
[Heaviside function](https://en.wikipedia.org/wiki/Heaviside_step_function)
```{r, echo=TRUE, eval=TRUE, cache=TRUE}
(y = ifelse(z > u, 1.0, 0.0))
```

# Perceptrons
The [Wikipedia](https://en.wikipedia.org/wiki/Perceptron) entry is a good place to start.

